\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

This thesis represents the culmination of a long journey, one that would have been impossible to complete without the unwavering support, encouragement, and guidance from the remarkable people around me. I am profoundly grateful to all who played a part in this process and contributed to this achievement.

First and foremost, I would like to express my deepest and most sincere gratitude to my thesis advisors. Your expert guidance, patience, and insightful feedback have been instrumental throughout this endeavor. Your belief in my potential pushed me to fully express myself, and for that, I am eternally thankful.

To my parents, no words can capture the depth of my gratitude. Thank you for always believing in my abilities and for the countless sacrifices you made to ensure I had every opportunity to succeed. Your unwavering support has been the foundation upon which my confidence and resilience are built. I owe everything to your love and faith in me.

To my brother, thank you for being the dual entity of my life. You have provided me with a different perspective on everything, helping me navigate challenges and see my strengths and weaknesses more clearly. Your presence has been a vital and irreplaceable part of this journey, bringing balance and clarity to my life when I needed it most.

To my closest friends, your loyalty and unwavering support over the years have been invaluable. Your friendship has been a continuous source of joy, laughter, shared experiences and constant growth. Special thanks to those who have been by my side since childhood and adolescence. You have been my pillars through thick and thin, and I am forever grateful for the memories we've created together.

To my university colleagues, thank you for making this academic journey so memorable. The late-night discussions, collaborative problem-solving sessions, shared triumphs and struggles have enriched this experience in ways I will cherish forever.

Finally, to everyone who has contributed to my personal and academic growth along the way - whether through direct support or quiet encouragement - thank you. This achievement is not mine alone; it stands as a testament to your generosity, wisdom and your enduring influence on my life.